---
title: "Comic Gender Disparity"
output: html_notebook
---


We first had to call specific libraries in order to reverse engineer our data. We used lubridate to help clean the data, tidyverse and refinr to organize it, and ggplot2/ggthemes to display it. We also had to create our libraries for the data we'd be referencing. We did so by loading the data into a data folder and then storing it in its respective variables.
```{r}
options(scipen=999)
library(lubridate)
library(tidyverse)
library(refinr)
library(ggplot2)
library(ggthemes)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 2)
```
```{r}

library(readr)
dc_wikia_data <- read_csv("data/dc-wikia-data.csv")
marvel_wikia_data <- read_csv("data/marvel-wikia-data.csv")


```


FiveThirtyEight:
"The same trend bears out when we look at characters with higher appearance counts. When we zero in on the 2,415 DC and 3,342 Marvel characters with gender data who appear at least 10 times — the set of characters that probably gives the most consistent view of what we see on the page — female characters make up only 30.9 percent of the DC universe and 30.6 percent of the Marvel universe."

"When we focus on only the most frequently appearing characters in each universe — the stars of the show — the data changes. Of characters with gender data and 100 or more appearances (294 DC characters and 414 Marvel characters), only 29.0 percent of DC’s are female, and a similar 31.1 percent of the Marvel crowd is."


Elie:
My two datasets were basically identical, with the only difference being the usage of the number 100 instead of the number 100 as a filter.  I began by applying the number filter, and then applying the gender filter.  I also created a new dataframe that created our percentages.  I did this by dividing our number of females seen over 10 or 100 times, and then multiplying that number by 100.  To make things look nicer, I also created a bar plot with male and female ratios.  I did this system for both Marvel and D.C.  characters.

```{r}
#Elie-Percent Female (10 >= appearances)

filter1_ten<-filter(dc_wikia_data, APPEARANCES >= 10) 
filter2_ten<-filter(filter1_ten, SEX != "NA") 


total_dc_ten=nrow(filter2_ten)
female_gender_count_ten<-nrow(filter(filter2_ten, SEX=='Female Characters'))

percentage_of_women_in_dc_ten<-(female_gender_count_ten/total_dc_ten)*100
print(percentage_of_women_in_dc_ten)


barplot(c(percentage_of_women_in_dc_ten, 100-percentage_of_women_in_dc_ten),names.arg=c("Female Ratio", "Male Ratio"), col="light green", main="DC Gender Ratio", xlab="Gender", ylab="Ratio")


filter1_ten<-filter(marvel_wikia_data, APPEARANCES >= 10)
filter2_ten<-filter(filter1_ten, SEX != "NA")


total_marvel_ten=nrow(filter2_ten)
female_gender_count_ten<-nrow(filter(filter2_ten, SEX=='Female Characters'))

percentage_of_women_in_marvel_ten<-(female_gender_count_ten/total_marvel_ten)*100

print(percentage_of_women_in_marvel_ten)


barplot(c(percentage_of_women_in_marvel_ten, 100-percentage_of_women_in_marvel_ten),names.arg=c("Female Ratio", "Male Ratio"), col="light green", main="Marvel Gender Ratio", xlab="Gender", ylab="Ratio")
```


```{r}
#Elie-Percent Female (100 >= appearances)

filter1<-filter(dc_wikia_data, APPEARANCES >= 100) 
filter2<-filter(filter1, SEX != "NA") 


total_dc=nrow(filter2)
female_gender_count<-nrow(filter(filter2, SEX=='Female Characters'))

percentage_of_women_in_dc<-(female_gender_count/total_dc)*100
print(percentage_of_women_in_dc)


barplot(c(percentage_of_women_in_dc, 100-percentage_of_women_in_dc),names.arg=c("Female Ratio", "Male Ratio"), col="light green", main="DC Gender Ratio", xlab="Gender", ylab="Ratio")


filter1<-filter(marvel_wikia_data, APPEARANCES >= 100)
filter2<-filter(filter1, SEX != "NA")


total_marvel=nrow(filter2)
female_gender_count<-nrow(filter(filter2, SEX=='Female Characters'))

percentage_of_women_in_marvel<-(female_gender_count/total_marvel)*100

print(percentage_of_women_in_marvel)


barplot(c(percentage_of_women_in_marvel, 100-percentage_of_women_in_marvel),names.arg=c("Female Ratio", "Male Ratio"), col="light green", main="Marvel Gender Ratio", xlab="Gender", ylab="Ratio")

```

FiveThirtyEight:
“Ms. Marvel is one of just 3,599 female characters in Marvel’s universe. To arrive at that number, I pulled data from the Marvel and DC Wikia databases.4 The sites are fan-maintained, so as with any data of this nature, take it with the standard caveat that it’s only as good as the people updating it.5 Still, these two sites are the best resources for this kind of information.”


JR:
The first thing I needed to do was filter for female characters so that Rosa and I could tackle the next problem. In the original article, this data was used to talk about Ms. or Captain Marvel who had recently starred in her own movie. I initially had a different result but was eventually able to figure out I needed to remove/filter for characters whose ID was "NA," in order to get the same results as FiveThirtyEight.
```{r}
#Jerold-Female Marvel Characters
marvel_female_characters <- marvel_wikia_data %>%
  filter(SEX == "Female Characters") %>%
  filter (ID != "NA") %>%
  summarize(
    count=n()
  )
```

FiveThirtyEight:
“Women and men were almost exactly equally likely to have a secret identity in the Marvel universe (49.4 percent of males and 49.5 percent of females), while in the DC universe 51 percent of males had a secret identity and 45 percent of females did. While it’s not a perfect stand-in, we can also infer that men may be slightly more likely to be superheroes or villains — rather than just normal, unpowered side characters — than women in the DC universe.”


JR/Rosa:
Rosa and I ended up working on this section together. It helped because my section was very easy and overlapped with her's. We used a lot of variables, more than needed, so that we could make sure we were getting good, accurate, results along the way. First we filtered for the men and women in each universe. Then we found the number of all four categories who had secret identities and determined the percentage in each category. We ended up getting the exact same results as FiveThirtyEight after we finished. After we knew we had the correct answer we wanted to display it all together and created the dc_percent and marvel_percent variables that showed men, women, and percentage at once for their respective universes.
```{r}
#Rosa-Secret Identities

dc_men <- dc_wikia_data %>%
  filter(SEX == "Male Characters") %>%
  filter (ID != "NA")

dc_men_identities <-
dc_wikia_data %>%
  filter(
    SEX == "Male Characters" & ID == "Secret Identity"
  )
dc_men_percent <-
  dc_men_identities %>%
  summarize(
    Gender = "Men",
    count_identities = n(),
    percentage_identity = (count_identities / 3422)*100
  )

dc_women <- dc_wikia_data %>%
  filter(SEX == "Female Characters") %>%
  filter (ID != "NA")

dc_women_identities <-
dc_wikia_data %>%
  filter(
    SEX == "Female Characters" & ID == "Secret Identity"
  )
dc_women_percent <-
  dc_women_identities %>%
  summarize(
    Gender = "Women",
    count_identities = n(),
    percentage_identity = (count_identities / 1390)*100
    )

marvel_men <- marvel_wikia_data %>%
  filter(SEX == "Male Characters") %>%
  filter (ID != "NA")

marvel_men_identities <-
marvel_wikia_data %>%
  filter(
    SEX == "Male Characters" & ID == "Secret Identity"
  )
marvel_men_percent <-
  marvel_men_identities %>%
  summarize(
    Gender = "Men",
    count_identities = n(),
    percentage_identity = (count_identities / 9128)*100
    )
    
marvel_women <- marvel_wikia_data %>%
  filter(SEX == "Female Characters") %>%
  filter (ID != "NA")

marvel_women_identities <-
marvel_wikia_data %>%
  filter(
    SEX == "Female Characters" & ID == "Secret Identity"
  )
marvel_women_percent <-
  marvel_women_identities %>%
  summarize(
    Gender = "Women",
    count_identities = n(),
    percentage_identity = (count_identities / 3047) *100
  )
```
```{r}
dc_percent <-
  bind_rows(dc_women_percent, dc_men_percent)
marvel_percent <-
  bind_rows(marvel_women_percent, marvel_men_percent)
```

FiveThirtyEight:
“I pulled a list of characters6 from the two mainstream universes of each publisher.7 This left me with 6,896 unique characters from DC and 16,376 unique characters from Marvel.”

Tatyana:
I ran into some issues trying to display this, so I kept my code for male and female characters separate. At first, I tried to group by align and sex but it wasn’t answering my question, which was: what is the alignment breakdown by gender? Once I started filtering for female characters and male characters, my tibbles were answering my question. I was able to come to the same conclusion as Hickey, which found that there are overwhelmingly more male characters than female characters. 


```{r}
#Tatyana-Alignment by Gender
dc_female_align <- dc_wikia_data %>% 
  filter(SEX=="Female Characters", ID !="NA") %>% 
  group_by(ALIGN) %>% 
  summarise(
    Gender = "Female",
    count_ALIGN =n(),
    Rate= (count_ALIGN/1390)*100
    )
dc_male_align <- dc_wikia_data %>% 
  filter(SEX=="Male Characters", ID !="NA") %>% 
  group_by(ALIGN) %>% 
  summarise(
    Gender = "Male",
    count_ALIGN =n(),
    Rate= (count_ALIGN/3422)*100
  )

marvel_female_align <- marvel_wikia_data %>% 
  filter(SEX=="Female Characters", ID !="NA") %>% 
  group_by(ALIGN) %>% 
  summarise(
    Gender = "Female",
    count_ALIGN =n(),
    Rate= (count_ALIGN/3047)*100
  )

marvel_male_align <- marvel_wikia_data %>% 
  filter(SEX=="Male Characters", ID !="NA") %>% 
  group_by(ALIGN) %>% 
  summarise(
    Gender = "Male",
    count_ALIGN =n(),
    Rate= (count_ALIGN/9128)*100
  )

```
```{r}
dc_align <- dc_female_align %>%
  left_join(dc_male_align, by="ALIGN")

marvel_align <- marvel_female_align %>%
  left_join(marvel_male_align, by="ALIGN")
```

