---
title: "lab_03"
author: "sean mussenden"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse here
library(tidyverse)
# Load janitor here
library(janitor)
```

## Load data
```{r}
# Write code to load the WV slice of the ppp loan data here and call it wv_ppp_loans.
# The data is called lab_03.rds, and is in the same folder as lab_03.Rmd (this data notebook)

wv_ppp_loans <- read_rds("lab_03.rds")

```

## Answer questions

Q1: Which five cities located in Kanawha County, West Virginia got the most loans (in number of loans)? How many did each get? To answer this question, you'll have to clean the very messy "city" field.

Note: To answer this question, please fully clean the city field for Kanawha County cities. Hint: once cleaned, the dataframe that counts the number of loans by city should have 38 distinct cities for Kanawha County.

A1:
```{r}

# Write your code in this codeblock.

kanawha <- wv_ppp_loans %>%
  filter(project_county_name == "KANAWHA")

kanawha_city <- kanawha %>%
  select(city) %>%
    mutate(city = str_to_upper(city)) %>%
  mutate(city=case_when(
    city == "BALLE" ~ "BELLE",
    city == "CHARELSTON" ~ "CHARLESTON",
    city == "CROSSLANES" ~ "CROSS LANES",
    city == "ELKVEIW" ~ "ELKVIEW",
    city == "S CHARLESTON" ~ "SOUTH CHARLESTON",
    city == "ST ALBANS" ~ "SAINT ALBANS",
    city == "ST. ALBANS" ~ "SAINT ALBANS",
    city == "SO CHARLESTON" ~ "SOUTH CHARLESTON",
    city == "ST.ALBANS" ~ "SAINT ALBANS",
    TRUE ~ city
    )) %>%
  group_by(city) %>%
  summarize(total_loans = n()) %>%
  arrange(desc(total_loans))
```

Q2: What is the name of the business that got the largest loan (by dollar value) in the 5-digit ZIP code in West Virginia that had the highest dollar amount in total loans (by dollar value)? Write up a short paragraph that also answers these questions:

* Where in West Virginia is this ZIP code located?
* How much was the business' loan for?
* How much of that loan was forgiven?
* Which number is bigger, the loan amount or the forgiveness amount? Can you explain the discrepancy?

A2: 
The West Virginia zipcode with the highest dollar amount taken in PPP loans was 25301. It is the zipcode for Charleston, WV. The company with the largest loan in Charleston was Jackson Kelly PLLC. There loan amount was $5,068,472.0 and the forgiveness amount is $5,126,618.60 which is more than the loan. This could be because of interest accrued since it was approved April 8, 2020 and wasn't forgiven until June 6, 2021.

```{r}
# Hint
# First, create a dataframe that shows the top 5-digit ZIP code by total dollar amount
zip5_wv <- wv_ppp_loans %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))
# Second, use the top ZIP code to create a separate dataframe that shows the business with the biggest loan in that ZIP code.
zip5_wv %>%
  group_by(zip) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))

wv_top_zip <- zip5_wv %>%
  filter(zip == "25301") %>%
  arrange(desc(amount))

```

Q3: Five business entities owned by the spouse of a member of West Virginia's U.S. House delegation got PPP loans. Write up a paragraph that answers the following questions.   

  * What are the businesses' names?
  * What do the businesses do?
  * How much money did the five businesses receive in total?
  * What percentage of the total loan amount have been forgiven?
  * How did the U.S. House member vote on the legislation initially authorizing the PPP loan program.

Feel free to add any interesting "color" (little interesting details) you find in your research.  

Some hints to help you answer this question:

* West Virginia's three U.S. House members are listed [here](https://www.govtrack.us/congress/members/WV#representatives)
* To find businesses they or their families own, you can search for financial disclosure forms for U.S. House members [here](https://disclosures-clerk.house.gov/PublicDisclosure/FinancialDisclosure#Search).
  * To speed things up, here are direct links to the three U.S. House members most recent financial disclosures, which I found by searching the site linked above.
  * [Rep. 1](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042042.pdf)
  * [Rep. 2](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10040330.pdf)
  * [Rep. 3](https://disclosures-clerk.house.gov/public_disc/financial-pdfs/2020/10042734.pdf)
* When you examine the disclosure forms, there may be lots of stock sales and retirement accounts mixed in.  Ignore those. In the table in the disclosure forms, look for "Partnership" or "C-Corp" in the column named "Income Type"
* This is a link to details on passage of the [CARES ACT](https://www.congress.gov/bill/116th-congress/house-bill/748). You can find a link to how the House voted by clicking on a link from [this page](https://www.congress.gov/bill/116th-congress/house-bill/748/actions).

A3:
The Representative is Carol Miller. Her husband's businesses names are DM MOTOR, INC, DUTCH MILLER OF CHARLESTON, INC, DUTCH MILLER OF CHARLOTTE, INC, DUTCH MILLER SUBARU, INC, and DUTCH MILLER CHEVROLET, INC. They are car dealerships. In total the businesses recieved $3,105,300. 101.1326% of the loan amounts have been forgiven. She voted for the legislation.
```{r}

# Hint
# First, create a dataframe that only has loans to those five companies
dutch_miller <- wv_ppp_loans %>%
  filter(str_detect(name, "DUTCH |DM MOTOR"))
# Second, use that newly created dataframe to calculate total loan amount and percentage forgiven for those companies
dm_total <- dutch_miller %>%
  summarize(
    total_amount = sum(amount),
    total_forgiveness_amount = sum(forgiveness_amount)
  ) %>%
  mutate(percent = (total_forgiveness_amount/total_amount)*100)
```
