---
title: "lab_01"
author: "Jerold Massie"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**). 
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**). 
* write out the answer in the form of a complete sentence in the space given (look for **A**). 
When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS. 

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse.To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}


# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data 

We're going to load the slice of PPP loans for West Virginia.

**Task** Run the codeblock below to load the data. 

```{r}
wv_ppp_loans <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data. 
**Q1** How many rows and columns does this dataframe have?
**A1** [write your answer here]

```{r}
glimpse(wv_ppp_loans)
```

## Answer questions

**Q2** Which West Virginia county got the most loans? How many?
**A2** [KANAWHA/4332]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange(desc(count_loans))
```
**Q3** Which West Virginia county got the fewest loans? How many?
**A3** [WIRT/36]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n()
  ) %>%
  arrange((count_loans))
```
**Q4** What's your best educated guess to explain why the county identified in question 2 had a lot of loans and why the county identified in question 3 had only a few loans? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.  
**A4** [According to 2020 census data, Kanawha County has a populaiton of 178,124, where as Wirt County has a population of 5,821.]


**Q5** What West Virginia business got the state's largest PPP loan? How much was it for? What does the business do? How many jobs did the loan retain, according to the data?
**A5** [GREENBRIAR HOTEL CORPORATION got the largest PPP loan in West Virginia for $8.9 million dollors. The company retained 500 jobs.]

```{r}
# Write the code to produce your answer here
wv_ppp_loans %>%
  arrange(desc(amount))
```

**Q6** What West Virginia politician (with his family) owns the company identified in question 2? What's your source for that information? Hint: this will require web research, you won't be able to answer this purely from the data set.  You do not need to write any code for this question, so no codeblock is provided.  
**A6** [According to a 2019 ProPublica article, West Virginia Gov. Jim Justice is the owner of the Greenbrier resort and is also the richest man in the state.
https://www.propublica.org/article/west-virginia-greenbrier-governor-jim-justice-little-trump]

**Q7**. Produce a summary table that has the following information for West Virginia as a whole (the whole state, not by county):
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings. 

**A7** [The total number of loans is 33703, the total amount is over $2.6 billion, the average amount for loans is $77,338, the median amount is $20,832, the smallest loan is for $165, and the largest loan is for over $8.9 million. The large discrepancy between the mean and median is something to look into, meaning there may have been few very large loans that are pulling the average up. Also the smallest loan of $165 could be an interesting piece on a smaller business, trying to answer why they needed such a specifc, small amount.]
```{r}
# Write the code to produce your answer here
wv_summary <-
  wv_ppp_loans %>%
  summarise(
    count_of_loans = n(),
    total_loan_amount = sum(amount),
    average_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    smallest_loan = min(amount),
    largest_loan = max(amount)
    
  ) 
```

**Q8** Produce a summary table that has the following information for each West Virginia county:
* count of loans
* total amount of all loans
* the average loan amount
* the median loan amount
* the smallest loan amount
* the largest loan amount

Store it as a variable object called `wv_county_summary` so that you can examine it with the data viewer. 

What newsworthy nuggets or patterns jump out at you that you could imagine dropping into a news story or doing follow up reporting on? Write up a few sentence that summarizes one county. 

**A8** [Kanawha County tied for the smallest loan amount with another county at $165, but also has a fairly high largest loan at $5.5 million. A comparison between the businesses at these two extremes and their circumstances could be very interesting. Another interesting tidbit is that the counties with the most loans don't neccessarily have the highest amount disbursed. For example, Cabell County disbrused 1993 loans and almost $235 million, and Monongalia county disbursed 2564 loans but only $181 million.]
```{r}
# Write the code to produce your answer here
wv_county_summary <-
  wv_ppp_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_of_loans = n(),
    total_loan_amount = sum(amount),
    average_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    smallest_loan = min(amount),
    largest_loan = max(amount)
    )
```
